The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free eLearning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2009-07-06 script/en/chapters/introduction.tex
[i18n] CT script: Cover, introduction and prime records updated

% --------------------------------------------------------------------------
\chapter*{Preface to the 8th Edition of the CrypTool Script\footnote{For the 9th edition no extra preface was written.}}
\addcontentsline{toc}{chapter}{Preface to the 8th Edition of the CrypTool Script}

@@

\vspace{12pt}
The rapid spread of the Internet has also lead to intensified research in the
technologies involved, especially within the area of cryptography where a good
deal of new knowledge has arisen.

%This edition of the script adds some topics, but mainly updates areas (e.g. the
%summaries of topical research areas):
Again this edition of the script amended and updated some topics, e.g.:
\vspace{-7pt}
\begin{itemize}
  \item the search for the largest prime numbers
        (Mersenne and Fermat primes, ``M44'' or the biggest non-Mersenne prime
         found in March 2007 when investigating Sierpinski numbers) \\ 
	(chap. \ref{spezialzahlentypen}, \ref{RecordPrimes};
         table \ref{L_n_Largest_Known-Primes}, 
         page \pageref{L_n_Largest_Known-Primes}), 
  \item the factorisation of big numbers like ``C307''\index{C307} 
        (chap. \ref{C307}),
  \item the current record (April 2007) calculating the discrete logarithm 
        (The used sample power consisted of the first 160 digits of Euler's
        number e and the calculation was done modulo a 160 digit prime)
        (chap. \ref{L_Discrete_Logarithm}),
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}).
\end{itemize}
Newly added is chapter \ref{Chapter_Crypto2020} {\bf Crypto 2020}
by Prof. Buchman et al. and the appendix \ref{s:appendix-Learn-NT} 
about the {\bf learning tool for number theory}.

\vspace{8pt}
The first time the document was delivered with CrypTool\index{CrypTool} 
was in version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool (1.2.02, 1.3.00, 1.3.02, 1.3.03, 1.3.04, 1.4.00
and now 1.4.10).

I'd be more than happy if this also continues in the further open-source
versions of CrypTool\index{CrypTool}.

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\[1.5\baselineskip]
\enlargethispage*{2\baselineskip}
\nopagebreak
Bernhard Esslinger
\\[\baselineskip]
Frankfurt (Germany), June 2007
\vspace{-12pt}

@@

CrypTool\index{CrypTool} is a program with an extremely comprehensive online
help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface.\par \vskip + 3pt

CrypTool\index{CrypTool} was developed during the end-user awareness program
at Deutsche Bank in order to increase employee awareness of IT security and provide them with
a deeper understanding of the term security.
A further aim has been to enable users to understand the cryptographic
procedures. In this way, using CrypTool
as a reliable reference implementation of the various encryption procedures,
you can test the encryption implemented in other programs. \par \vskip + 3pt

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. Without their special
talents and engagement CrypTool would not be what it is today:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann and Prof.\ Torben Weis.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\

@@

Frankfurt (Germany), July 2008

>>>>

% --------------------------------------------------------------------------
\chapter*{Preface to the 10th Edition of the CrypTool Script}
\addcontentsline{toc}{chapter}{Preface to the 10th Edition of the CrypTool Script}

@@

The rapid spread of the Internet has also lead to intensified research in the
technologies involved, especially within the area of cryptography where a good
deal of new knowledge has arisen.

%This edition of the script adds some topics, but mainly updates areas (e.g. the
%summaries of topical research areas):
Again this edition of the script corrected, amended and updated some topics, e.g.:
\vspace{-7pt}
\begin{itemize}
  \item the search for the largest prime numbers, 
%  \item progress in cryptanalysis of hash algorithms 
%        (chap. \ref{collision-attacks-against-sha-1}) and
%  \item progress in ideas for new crypto methods (RSA successor) 
%        (chap. \ref{xxxxxxxxxBrute-force-gegen-Symmetr})\index{xxxxxxxxxxxxxxxx} and
  \item the list of movies or novels, in which cryptography or number theory 
        played major role (see appendix \ref{s:appendix-movies});
        and where primes are used as hangers  
        (see curiouses in \ref{HT-Quaint-curious-Primes-usage}).
\end{itemize}

Newly added is the appendix about using the computer algebra system
Sage \ref{s:appendix-using-sage}. This was added because Sage becomes
more and more the standard open-source CAS system. Accordingly all
samples written before in PARI-GP and Mathematica have been substituted
with Sage code. Thanks to Minh Van Nguyen a lot of new code samples could be added.

\vspace{8pt}
The first time the document was delivered with CrypTool\index{CrypTool} 
was in version 1.2.01. Since then it has been expanded and revised in almost
every new version of CrypTool.

@@

I hope that many readers have fun with this script and that they get 
out of it more interest and greater understanding of this modern but 
also very ancient topic.
\\[1.5\baselineskip]
\enlargethispage*{2\baselineskip}
\nopagebreak
Bernhard Esslinger
\\[\baselineskip]
Frankfurt (Germany), July 2009
\vspace{-12pt}

@@

CrypTool\index{CrypTool} is an educational program with a comprehensive online
help enabling you to use and analyse cryptographic procedures within a
unified graphical user interface.\par \vskip + 3pt

@@

At this point I'd like to thank explicitly the following people who
particularly contributed to CrypTool\index{CrypTool}. They applied their
very special talents and showed really great engagement:
\vspace{-7pt}
%\begin{itemize}
\begin{list}{\textbullet}{\addtolength{\itemsep}{-0.5\baselineskip}}
   \item Mr.\ Henrik Koy
   \item Mr.\ J\"org-Cornelius Schneider
   \item Mr.\ Florian Marchal
   \item Dr.\ Peer Wichmann
   \item Staff of Prof.\ Claudia Eckert, Prof.\ Johannes Buchmann and Prof.\ Torben Weis.
\end{list}
%\end{itemize}
Also I want to thank all the many people not mentioned here for their 
hard work (mostly carried out in their spare time).
\\

@@

Frankfurt (Germany), July 2009

### 2009-07-06 script/en/chapters/primes.tex
[i18n] CT script: Cover, introduction and prime records updated

\newpage
\hypertarget{Kapitel_2}{}
\chapter{Prime Numbers}
\label{Label_Kapitel_2}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, June 2007)

@@

% --------------------------------------------------------------------------
\hypertarget{RecordPrimes}{}
\subsection{The 20 largest known primes (as of May 2007)}  % Eyecatcher_neue-Mersenne
\label{RecordPrimes}
\index{Prime number!records}

The following table contains the current record primes and
a description of its particular number type\footnote{%
An up-to-date version can be found in the internet at
     \href{http://primes.utm.edu/largest.html}
  {\texttt{http://primes.utm.edu/largest.html}}.
}:

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|cccc|}
\hline    % Eyecatcher_neue-Mersenne
	& {\bf Definition} & {\bf Decimal Digits} & {\bf When} & {\bf Description} \\
\hline
	1  & $2^{32,582,657}-1$ & 9,808,358 & 2006 & Mersenne, 44th known \\
	2  & $2^{30,402,457}-1$ & 9,152,052 & 2005 & Mersenne, 43rd known \\
	3  & $2^{25,964,951}-1$ & 7,816,230 & 2005 & Mersenne, 42nd known \\
	4  & $2^{24,036,583}-1$ & 7,235,733 & 2004 & Mersenne, 41st known \\
	5  & $2^{20,996,011}-1$ & 6,320,430 & 2003 & Mersenne, 40th known \\
	6  & $2^{13,466,917}-1$ & 4,053,946 & 2001 & Mersenne, M-39 \\
	7  & $19,249 \cdot 2^{13,018,586}+1$ & 3,918,990 & 2007 & Generalized Mersenne\footnotemark \\
	8  & $27,653 \cdot 2^{9,167,433}+1$ & 2,759,677 & 2005 & Generalized Mersenne \\

	9  & $28,433 \cdot 2^{7,830,457}+1$ & 2,357,207 & 2004 & Generalized Mersenne \\

	10  & $2^{ 6,972,593}-1$ & 2,098,960 & 1999 & Mersenne, M-38 \\
	11  & $5,359 \cdot 2^{5,054,502}+1$ & 1,521,561 & 2003 & Generalized Mersenne \\

	12  & $4,847 \cdot 2^{3,321,063}+1$ & 999,744 & 2005 & Generalized Mersenne \\
	13  & $3 \cdot 2^{3,136,255}-1$ & 944,108 & 2007 & Generalized Mersenne \\

	14  & $2^{ 3,021,377}-1$ &   909,526 & 1998 & Mersenne, M-37 \\
	15  & $2^{ 2,976,221}-1$ &   895,932 & 1997 & Mersenne, M-36 \\

	16  & $222,361 \cdot 2^{2,854,840}+1$ & 859,398 & 2006 & Generalized Mersenne \\

	17 & $1,372,930^{131,072}+1$ &   804,474 & 2003 & Generalized Fermat\footnotemark \\

	18  & $1,361,244^{131,072}+1$ &   803,988 & 2004 & Generalized Fermat \\
	19  & $1,176,694^{131,072}+1$ &   795,695 & 2003 & Generalized Fermat \\

	20  & $342,673 \cdot 2^{2,639,439}-1$ & 794,556 & 2007 & Generalized Mersenne \\

\hline
\end{tabular}
\caption{The 20 largest known primes and its particular number types
         (as of May 2007)}    % Eyecatcher_neue-Mersenne
\label{L_n_Largest_Known-Primes}
\end{center}
\end{table} 
\footnotetext{\index{Number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB) (\href{http://www.seventeenorbust.com}{\texttt{http://www.seventeenorbust.com}})
% {\href{http://www.mersenne.org} {\tt http://www.mersenne.org}}
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its
task sometime.

@@

Due to the fact that the exponents of the Mersenne numbers
\index{Mersenne!number} do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers 
\index{Mersenne!prime number} have the exponents\footnote{%
The following page from Landon Curt Noll\index{Noll, Landon Curt} contains 
all Mersenne primes including its date of discovery and its value as number 
and as word:
      \href{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}
   {\texttt{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}}. \\
Also see:
      \href{http://www.utm.edu/}
   {\texttt{http://www.utm.edu/}}.
                             } 
$$
\begin{array}{c}
2; ~ 3; ~ 5; ~ 7; ~ 13; ~ 17; ~ 19; ~ 31; ~ 61; ~ 89; ~ 107; ~ 127;
~ 521; ~ 607; ~ 1,279; ~ 2,203; ~ 2,281; ~ 3,217; ~ 4,253;\\
 4,423; ~ 9,689; ~ 9,941, ~ 11,213; ~ 19,937; ~ 21,701; ~ 23,207; ~ 44,497; ~
86,243; ~ 110,503; ~ 132,049; \\
 216,091; ~ 756,839; ~ 859,433; ~ 1,257,787; ~ 1,398,269; ~ 2,976,221;
 ~ 3,021,377; ~ 6,972,593; \\
 ~ 13,466,917;  ~ 20,996,011;  ~ 24,036,583;  ~ 25,964,951;  ~ 30,402,457,  ~ 32.582.657.
% be_2005_UPDATEN_if-new-mersenne-prime-appears          ~ xxx,xxx,xxx.
\end{array}
$$
Thus    % Eyecatcher_neue-Mersenne
$44$            % be_2005_UPDATEN_if-new-mersenne-prime-appears
Mersenne prime numbers are currently known%
\index{Prime number!Mersenne}\index{Mersenne!prime number}.

@@

Right now (May 2005) all prime exponents smaller than $ 13.466.917 $ have been
tested and double-checked (see home page of the GIMPS project:
{\href{http://www.mersenne.org} {\tt http://www.mersenne.org}}): so we can
be certain, that this is really the 39th Mersenne prime number and that
there are no smaller undiscovered Mersenne primes (it is common usage to use
the notation M-nn not until it is proven, that the nn-th known Mersenne prime
is really the nn-th Mersenne prime).

@@

Right now the GIMPS project has discovered ten
   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
largest Mersenne primes so far, including the largest known prime number at all.

@@

   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
\begin{table}[h]
\begin{center}
\begin{tabular}{|cccc|}
\hline
	{\bf Definition} & {\bf Decimal Digits} & {\bf When} & {\bf Who} \\
\hline
	$2^{32,582,657}-1$ & 9,808,358 & September 4, 2006 & Curtis Cooper/Steven Boone \\
	$2^{30,402,457}-1$ & 9,152,052 & December 15, 2005 & Curtis Cooper/Steven Boone \\
	$2^{25,964,951}-1$ & 7,816,230 & February 18, 2005 & Martin Nowak     \\
	$2^{24,036,583}-1$ & 7,235,733 & May 15, 2004      & Josh Findley     \\
	$2^{20,996,011}-1$ & 6,320,430 & November 17, 2003 & Michael Shafer   \\
	$2^{13,466,917}-1$ & 4,053,946 & November 14, 2001 & Michael Cameron  \\
	$2^{ 6,972,593}-1$ & 2,098,960 & June 1, 1999      & Nayan Hajratwala \\
	$2^{ 3,021,377}-1$ &   909,526 & January 27, 1998  & Roland Clarkson  \\
	$2^{ 2,976,221}-1$ &   895,932 & August 24, 1997   & Gordon Spence    \\
	$2^{ 1,398,269}-1$ &   420,921 & November 1996     & Joel Armengaud   \\

\hline
\end{tabular}
   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
\caption{The largest primes found by the GIMPS project (as of May 2007)}
\end{center}
\end{table}

@@

% --------------------------------------------------------------------------
\vskip +25 pt
\subsection{Challenge of the Electronic Frontier Foundation (EFF)}\index{EFF}
This search is also spurred on by a competition started by the non-profit
organisation EFF (Electronic Frontier Foundation) using the means of an 
unknown donor. The participants are rewarded with a total of 500,000 USD if
they find the longest prime number. In promoting this project, the unknown
donor is not looking for the quickest computer, but rather wants to draw 
people's attention to the opportunities offered by {\em cooperative networking} \\
{\href{http://www.eff.org/coopawards/prime-release1.html}{\tt http://www.eff.org/coopawards/prime-release1.html}}

@@

The next prize of 100,000 USD offered by EFF is for a proven prime with more
than 10 million decimal digits.
%{\href{http://www.octocad.demon.co.uk/mersenne/prime.htm }{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm}}.

>>>>

\newpage
\hypertarget{Kapitel_2}{}
\chapter{Prime Numbers}
\label{Label_Kapitel_2}
(Bernhard Esslinger, May 1999; Updates Nov. 2000, Dec. 2001, June 2003, May 2005, March 2006, June 2007, July 2009)

@@

% --------------------------------------------------------------------------
\hypertarget{RecordPrimes}{}
\subsection{The 20+ largest known primes (as of July 2009)}  % Eyecatcher_neue-Mersenne
\label{RecordPrimes}
\index{Prime number!records}

The following table contains the current record primes and
a description of its particular number type\footnote{%
An up-to-date version can be found in the internet at
     \href{http://primes.utm.edu/largest.html}
  {\texttt{http://primes.utm.edu/largest.html}}
and
     \href{http://primes.utm.edu/mersenne/index.html}
  {\texttt{http://primes.utm.edu/mersenne/index.html}}.
}:

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|cccc|}
\hline    % Eyecatcher_neue-Mersenne
	& {\bf Definition} & {\bf Decimal Digits} & {\bf When} & {\bf Description} \\
\hline
	1  & $2^{43,112,609}-1$ & 12,978,189 & 2008 & Mersenne, 47th known \\
	2  & $2^{42,643,801}-1$ & 12,837,064 & 2009 & Mersenne, 46rd known \\
	3  & $2^{37,156,667}-1$ & 11,185,272 & 2008 & Mersenne, 45nd known \\

	4  & $2^{32,582,657}-1$ &  9,808,358 & 2006 & Mersenne, 44th known \\
	5  & $2^{30,402,457}-1$ &  9,152,052 & 2005 & Mersenne, 43rd known \\
	6  & $2^{25,964,951}-1$ &  7,816,230 & 2005 & Mersenne, 42nd known \\
	7  & $2^{24,036,583}-1$ &  7,235,733 & 2004 & Mersenne, 41st known \\
	8  & $2^{20,996,011}-1$ &  6,320,430 & 2003 & Mersenne, 40th known \\
	9  & $2^{13,466,917}-1$ &  4,053,946 & 2001 & Mersenne, M-39 \\
	10 & $19,249 \cdot 2^{13,018,586}+1$ & 3,918,990 & 2007 & Generalized Mersenne\footnotemark \\
	11 & $27,653 \cdot 2^{9,167,433}+1$ & 2,759,677 & 2005 & Generalized Mersenne \\

	12 & $28,433 \cdot 2^{7,830,457}+1$ & 2,357,207 & 2004 & Generalized Mersenne \\

	13  & $2^{ 6,972,593}-1$ & 2,098,960 & 1999 & Mersenne, M-38 \\
	14  & $5,359 \cdot 2^{5,054,502}+1$ & 1,521,561 & 2003 & Generalized Mersenne \\

	15  & $4,847 \cdot 2^{3,321,063}+1$ & 999,744 & 2005 & Generalized Mersenne \\
	16  & $3 \cdot 2^{3,136,255}-1$ & 944,108 & 2007 & Generalized Mersenne \\

	17  & $2^{ 3,021,377}-1$ &   909,526 & 1998 & Mersenne, M-37 \\
	18  & $2^{ 2,976,221}-1$ &   895,932 & 1997 & Mersenne, M-36 \\

	19  & $222,361 \cdot 2^{2,854,840}+1$ & 859,398 & 2006 & Generalized Mersenne \\

	20 & $1,372,930^{131,072}+1$ &   804,474 & 2003 & Generalized Fermat\footnotemark \\

	21  & $1,361,244^{131,072}+1$ &   803,988 & 2004 & Generalized Fermat \\
	22  & $1,176,694^{131,072}+1$ &   795,695 & 2003 & Generalized Fermat \\

	23  & $342,673 \cdot 2^{2,639,439}-1$ & 794,556 & 2007 & Generalized Mersenne \\

\hline
\end{tabular}
\caption{The 20+ largest known primes and its particular number types
         (as of July 2009)}    % Eyecatcher_neue-Mersenne
\label{L_n_Largest_Known-Primes}
\end{center}
\end{table} 
\footnotetext{\index{Number!Sierpinski}\index{Seventeen or Bust SoB}%
This number was found within the distributed computing project
``Seventeen or Bust'' (SoB) (\href{http://www.seventeenorbust.com}{\texttt{http://www.seventeenorbust.com}})
% {\href{http://www.mersenne.org} {\tt http://www.mersenne.org}}
at March 26, 2007. While the well known \hyperlink{GIMPS-project}{GIMPS project} (chapter~\ref{zahlentyp_mersenne}) searches for bigger and bigger of the infinitely many primes, there is a chance, that the SoB project could have been completed its
task sometime.

@@

Due to the fact that the exponents of the Mersenne numbers
\index{Mersenne!number} do not use all
natural numbers, but only the primes, the {\em experimental space} is limited
considerably. The currently known Mersenne prime numbers 
\index{Mersenne!prime number} have the exponents\footnote{%
The following page from Landon Curt Noll\index{Noll, Landon Curt} contains 
all Mersenne primes including its date of discovery and its value as number 
and as word:
      \href{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}
   {\texttt{http://www.isthe.com/chongo/tech/math/prime/mersenne.html}}. \\
Also see:
      \href{http://www.utm.edu/}
   {\texttt{http://www.utm.edu/}}.
                             } 
$$
\begin{array}{c}
2; ~ 3; ~ 5; ~ 7; ~ 13; ~ 17; ~ 19; ~ 31; ~ 61; ~ 89; ~ 107; ~ 127;
~ 521; ~ 607; ~ 1,279; ~ 2,203; ~ 2,281; ~ 3,217; ~ 4,253;\\
 4,423; ~ 9,689; ~ 9,941, ~ 11,213; ~ 19,937; ~ 21,701; ~ 23,207; ~ 44,497; ~
86,243; ~ 110,503; ~ 132,049; \\
 216,091; ~ 756,839; ~ 859,433; ~ 1,257,787; ~ 1,398,269; ~ 2,976,221;
 ~ 3,021,377; ~ 6,972,593; \\
 ~ 13,466,917; ~ 20,996,011; ~ 24,036,583; ~ 25,964,951; ~ 30,402,457; ~ 32,582,657;\\
 ~ 37,156,667; ~ 43,112,609; ~ 42,643,801.
% be_2005_UPDATEN_if-new-mersenne-prime-appears          ~ xxx,xxx,xxx.
\end{array}
$$
Thus    % Eyecatcher_neue-Mersenne
$47$            % be_2005_UPDATEN_if-new-mersenne-prime-appears
Mersenne prime numbers are currently known%
\index{Prime number!Mersenne}\index{Mersenne!prime number}.

@@

Right now (July 2009) all prime exponents smaller than $ 18,000,989 $ have been
tested and double-checked\footnote{%
See home page of the GIMPS project:
      \href{http://www.mersenne.org/report_milestones}
   {\texttt{http://www.mersenne.org/report\_milestones/}}.
                                  }:
so we can
be certain, that this is really the 39th Mersenne prime number and that
there are no smaller undiscovered Mersenne primes (it is common usage to use
the notation M-nn not until it is proven, that the nn-th known Mersenne prime
is really the nn-th Mersenne prime).

@@

Right now the GIMPS project has discovered 13
   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
largest Mersenne primes so far, including the largest known prime number at all.

@@

   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
\begin{table}[h]
\begin{center}
\begin{tabular}{|cccc|}
\hline
	{\bf Definition} & {\bf Decimal Digits} & {\bf When} & {\bf Who} \\
\hline

	$2^{43,112,609}-1$ & 12,978,189 & August 23, 2008 & Edson Smith  \\
	$2^{42,643,801}-1$ & 12,837,064 & April 12, 2009 & Odd Magnar Strindmo \\
	$2^{37,156,667}-1$ & 11,185,272 & September 6, 2008 & Hans-Michael Elvenich \\

	$2^{32,582,657}-1$ &  9,808,358 & September 4, 2006 & Curtis Cooper/Steven Boone \\
	$2^{30,402,457}-1$ &  9,152,052 & December 15, 2005 & Curtis Cooper/Steven Boone \\
	$2^{25,964,951}-1$ &  7,816,230 & February 18, 2005 & Martin Nowak     \\
	$2^{24,036,583}-1$ &  7,235,733 & May 15, 2004      & Josh Findley     \\
	$2^{20,996,011}-1$ &  6,320,430 & November 17, 2003 & Michael Shafer   \\
	$2^{13,466,917}-1$ &  4,053,946 & November 14, 2001 & Michael Cameron  \\
	$2^{ 6,972,593}-1$ &  2,098,960 & June 1, 1999      & Nayan Hajratwala \\
	$2^{ 3,021,377}-1$ &    909,526 & January 27, 1998  & Roland Clarkson  \\
	$2^{ 2,976,221}-1$ &    895,932 & August 24, 1997   & Gordon Spence    \\
	$2^{ 1,398,269}-1$ &    420,921 & November 1996     & Joel Armengaud   \\

\hline
\end{tabular}
   % be_2005_UPDATEN_if-new-mersenne-prime-appears    % Eyecatcher_neue-Mersenne
\caption{The largest primes found by the GIMPS project (as of July 2009)}
\end{center}
\end{table}

@@

% --------------------------------------------------------------------------
\vskip +25 pt
\subsection{Challenge of the Electronic Frontier Foundation (EFF)}\index{EFF}
This search is also spurred on by a competition started by the non-profit
organisation EFF (Electronic Frontier Foundation) using the means of an 
unknown donor. The participants are rewarded with a total of 500,000 USD if
they find the longest prime number. In promoting this project, the unknown
donor is not looking for the quickest computer, but rather wants to draw 
people's attention to the opportunities offered by {\em cooperative networking} \\
{\href{http://www.eff.org/awards/coop}{\tt http://www.eff.org/awards/coop}}

@@

% be_2005_UPDATEN_if-new-mersenne-prime-appears
For the next prize of 100,000 USD offered by EFF for a proven prime with more
than 10 million decimal digits, Edson Smith qualified, who found within the
GIMPS project $ 2^43,112,609-1 $.
%{\href{http://www.octocad.demon.co.uk/mersenne/prime.htm }{\tt http://www.octocad.demon.co.uk/mersenne/prime.htm}}.

### 2009-07-06 script/en/script-en.tex
[i18n] CT script: Cover, introduction and prime records updated

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{ 
	The CrypTool Script: \\
	Cryptography, Mathematics and More \\
        \vspace{30pt}
        {\large	Background reading\\%*[-8pt] 
        for the free eLearning program CrypTool\\
       (9th edition -- distributed with CrypTool version 1.4.20)}
        \vspace{60pt}
}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{ 
	The CrypTool Script: \\
	Cryptography, Mathematics and More \\
        \vspace{30pt}
        {\large	Background reading\\%*[-8pt] 
        for the free eLearning program CrypTool\\
       (10th edition -- distributed with CrypTool version 1.4.30)}
        \vspace{60pt}
}
